Investigating Neolithic pottery usage from Guijiabao in the Yanyuan Basin in Southwest China using organic residue analysis

### 2023-02-01 analysis/paper/paper.Rmd
worked on the conclusion section

Our investigation of the ceramics excavated from Southwest China suggests that a mixture of C3 plants and ruminant animals could be the food sources in this region during the Neolithic transition. This case study not only indicates the possibility of rice agriculture that has not been identified previously at the site but also demonstrates the first use of lipid analysis for pottery to explore the diet of the late Neolithic period in this region. Our case study not only indicates the possibility of rice agriculture that has not been identified previously at the site but also demonstrates the first use of lipid analysis for pottery to explore the diet of the late Neolithic period in this region.

The results show a mixture of C3 plants, such as rice, and terrestrial animals for most pots with one pot showing more C4 plant sources or marine sources. Our striking finding might suggest that the consumption of rice could be part of daily life at Guijiabao and the adoption of rice agriculture could be as important as millet cultivation which was thought the main crop previously.

More systematic sampling of potsherds across different phases and pottery types may eliminate a clear picture of diet and subsistence during the Neolithic transition.

>>>>

Our study on organic residues preserved in pottery fragments from Guijiabao and Gaoshan in southwest China reveals the presence of lipids but in a relatively low amount for most potsherds. The food cooked in those ceramics include ruminant fats and terrestrial plants during the Neolithic transition when the mixing farming of rice/millet spread into southwest China. It is most likely the terrestrial plant signal is C3 plant, implying rice processing in the pot. However, the evidence of direct processing of terrestrial plants is not as much as we expected according to the large quantity of plant remains, such as millet grains, found at Guijuabao. Instead, food processing in the early Neolithic is mostly ruminant fats that could be from local wild animals, such as deer. This might indicate that the proportion of rice/millet in the local diet is not high during the Neolithic transition, suggesting that hunting might be the primary subsistence, supplemented by rice/millet farming. This study offers a perspective from organic residue analysis to explore food choices and the extent of farming adoption when rice/millet agriculture first spread into the Yanyuan during the Neolithic period. We recognize the limitation of our interpretation of the local diet due to the small sample size. A more systematic sampling of potsherds across different phases and pottery types may eliminate a clear picture of diet and subsistence during the Neolithic transition.
